---
title: "TimeSeries_CaseStudy1"
author: "Vivek Sahoo"
date: "11/19/2019"
output:
  word_document: default
  html_document: default
---

The first step is to load the data, visualize the time series, and plot its ACF and PACF and only then identify the appropriate ARMA model to fit.

## 1) Data Visualization

```{r}
#Visualize Time Series, plot ACF and PACF
data=as.ts(scan("tsHW3.txt"))
par(mfrow=c(1,3))
plot(data)
acf(data)
pacf(data)
```

## (2) the identification of the model to be fit together with your rational based upon the appropriate plots and outputs

Observations from the above plots:

1) Time Series: It looks like a non-stationary time series which can be remedied by first order differencing or using an ARIMA model with the differncing parameter set to 1. But since we have been asked to assume that the model is stationary, we will refrain from using the difference operator.

2) ACF: It seems to be exponentially decreasing after lag 0. Hence we should set the order of MA term to 0.

3) PACF: The PACF seems to get cut off at a lag of 2. This gives us strong reason to set the order of AR term to 2.

So we will fit the model ARMA(2,0) and see if further revision is required.


## 3) estimation of the model parameters and interpretation of significance of the parameters

```{r}
fit = arima(x = data, order = c(2, 0, 0))
#Model parameters and significance of ARMA(2,0)
fit
```


### ARMA(2,0) or AR(2)
The Model Parameters ϕ1 = 0.484, ϕ2 = 0.3224 and intercept(mean) = 37981.4 can all be considered significant since the corresponding z values which can be obtained by dividing the parameter by its standard error are 5.01 for ϕ1, 3.263 for ϕ2 and 24.6 for intercept.

These z values are substantially greater than 1.96 which is the threshold to reject the null hypothesis that the paramters have a value of 0.


## (4) diagnostic checking on residuals and possible revision to your model choice

```{r}
#plotting residuals, ACF & PACF
par(mfrow=c(1,3))
plot(fit$resid)
acf(fit$resid)
pacf(fit$resid)
```

Plotting the residuals against time, we see the distribution might actually be white noise. But in order to confirm this, we also take ACF and PACF into consideration. From the ACF and PACF plots, we get strong evidence that the residual distribution is white noise, since none of the ACF & PACF values are above the significant level (except ACF at lag 0 which is always 1)

We also perform the 'Ljung-Box test' to test the null hypothesis that the model is a good fit to the data based on residuals.

```{r}
Box.test(resid(fit), lag = 3, type = "Ljung-Box", fitdf = 2)
```

Based on the p value for the above test, we can't reject the null hypothesis. Hence we can conclude our model is a good fit to the data.

Let's visualize our fitted model with the data.

```{r}
plot.ts(data)
points(fitted(fit),pch=20,col="red")
points(fitted(fit),type="l",col="red")
```

The Model seems to fit well with the data.

## (5) forecasts for the period in question (2 years)

```{r}
library(forecast)

plot(forecast(fit, h=24))
```


