---
title: "7050_FinalProject3_4"
author: "Vivek Sahoo"
date: "12/3/2019"
output:
  word_document: default
  html_document: default
---

#Question 3

```{r}
library(tseries)
data2 = read.csv("MORTGAGE30USsince2005.csv")
data2 = as.ts(data2[,2])
data3 = read.csv("MORTGAGE5USsince2005.csv")
data3 = as.ts(data3[,2])
data4 = read.csv("MORTGAGE15USsince2005.csv")
data4 = as.ts(data4[,2])


mge23 = as.matrix(cbind(data2,data3), demean = FALSE)
po.test(mge23)

mge34 = as.matrix(cbind(data3,data4), demean = FALSE)
po.test(mge34)

mge24 = as.matrix(cbind(data2,data4), demean = FALSE)
po.test(mge24)
```

#Analysing Results

15-Year Fixed Rate Mortgage and 5/1-Year Adjustable Rate Mortgage combined give the best results. But none of the pairs can reject the null hypothesis of non co-integration.

#Question 4

```{r}
cpi = read.csv("CPIAUCSLsince1954.csv")
cpi = as.ts(cpi[,2])
fund = read.csv("FEDFUNDSsince1954.csv")
fund = as.ts(fund[,2])
unemploy = read.csv("UNRATEsince1954.csv")
unemploy = as.ts(unemploy[,2])

econ = as.matrix(cbind(cpi,fund,unemploy), demean = FALSE)

library(vars)

VARselect(econ, lag.max=8,
          type="const")[["selection"]]
```

#Determining Lag Length
As per AIC we suggests the lag length of 5 whereas BIC(SC) suggets 3. 
Let's iteratively check the p value for Portmanteau test starting with 3.

```{r}
var3 <- VAR(econ, p=3, type="const")
serial.test(var3, lags.pt=10, type="PT.asymptotic")
```

It's suggests non stationarity. Let's try lag length of 4 and 5.

```{r}
#var4
var4 <- VAR(econ, p=4, type="const")
serial.test(var4, lags.pt=10, type="PT.asymptotic")

#var5
var5 <- VAR(econ, p=5, type="const")
serial.test(var5, lags.pt=10, type="PT.asymptotic")
```

We still have to reject the null hypothesis. We will choose var(3) which suggests the least complex model

```{r}
var3

#Granger Causality
summary(var3)
```

We can spot the significant lags per the output above

```{r}
library(forecast)
plot(forecast(var3, h=50))
```

There's slight upward trend as compared with the result in 2 since CPI has a high upward trend which influences the forecast